### 2023-11-09 main.tex
^^

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
% MACROS
% % 
% \newcommand{\cristian}[1]{{\bf (* {\color{ForestGreen}V:{} \small #1}*)}}
% \newcommand{\jorge}[1]{{\bf (* {\color{BlueViolet}G:{} \small #1}*)}}
% \newcommand{\rouven}[1]{{ (* {\color{red}{} \small #1}*)}}
 \newtheorem{remark}{Remark}
\frenchspacing

@@

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{u(x)}}
\newcommand{\oad}{\mathbf{a}^{\dagger}}

@@

	H=H_{left}+H_{bulk}+H_{right}

@@

	H_{bulk}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}

@@

	H_{right}=\begin{pmatrix}

@@

We define also 
\begin{equation}\label{lambdaConditions}
	\lambda_{a}=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}\lambda_{a}=0\,.

@@

the extra-site $u(x)$ connected to $x\in V$.

We define the duality function of the open multi-species stirring process as
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
\end{equation}
where we recall the notation (see \eqref{rhox}) for the \textit{density} of the species $a\in \{1,\ldots,N-1\}$ imposed by the reservoir at site $x\in V$:
\begin{equation}
\rho_{a}^{x}=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
\end{equation}
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
\end{equation}
where 
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
\end{equation}
\newline
Therefore the dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\newline \newline
\begin{remark} in the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigotimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\begin{equation}
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}\dt
\end{equation}
\end{remark}
\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{DualityRelation} is fulfilled.  To show this, we will use the Hamiltonians and their Lie algebraic description. Indeed, in this formalism the proof 
reduces to finding symmetries of the generator (for the bulk duality) and group like transformations (for the ``boundary duality'').

To a configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of a dual process we associate the vector
\begin{equation}
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle\right)\dt
\end{equation}
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
\end{equation}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{u(x)}\,E_{Na}^{x}-E_{aa}^{x}\right)\dt
\end{equation}
Here we introduced the pair of bosonic operators $\mathbf{a},\,\mathbf{a}^{+}$ satisfying $[\mathbf{a},\mathbf{a}^{+}]=1$ and acting as
\begin{equation}
	\mathbf{a}^{+}|q\rangle=|q+1\rangle\qquad \mathbf{a}|q\rangle=q|q-1\rangle
\end{equation}
\begin{equation}
	\langle q|\mathbf{a}^{+}=\langle q-1|\qquad \langle q|\mathbf{a}=(q+1) \langle q+1|
\end{equation}
on a generic vector $\langle q|$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(\mathbf{a}^{+})_{a}^{u(x)}$ denotes a copy of $\mathbf{a}^{+}$ acting on the extra-site $u(x)$ and on the species $a\in\{1,\ldots,N-1\}$. \\

@@

%\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N}^{u(x)}|(A^{\dagger})_{a}^{u(x)}=0

@@

\left\{d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\right\}\,|\alpha^{x}|\sum_{a=1}^{N-1}\left((A^{\dagger})_{a}^{u(x)}E_{Na}^{x}-E_{aa}^{x}\right)   

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Subsection \ref{subsection-description-process-LINE}. We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. Now we assume the hard-core exclusion rule, i.e. $\nu=1$. The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of $gl(N)$. Duality result stated in Subsection \ref{statementDualitySubsection} can be adapted. The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{left}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{right}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{left}=\sum_{a=1}^{N-1}\left((\mathbf{a}{\dagger})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{right}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{\dagger})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(a^{\dagger})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Sections \ref{sectionIntegrabiliy}.\\

@@

	\left[X_{a},X_{b}\right]=\lambda_{a}X_{b}-\lambda_{b}X_{a}\qquad\forall a,b=1,\ldots,N

@@

	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=\lambda_{a}\langle V^{*}|\qquad\forall a=1,\ldots,N\co

@@

	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-\lambda_{a}|V\rangle\qquad\forall a=1,\ldots,N\dt

@@

		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}\lambda_{a}\langle V^{*}|=0
	and using  \eqref{ratesConditions} and \eqref{lambdaConditions}.

@@

 The first transformation $S_{1}$ is closely related to the duality matrix and we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalizes the right boundary. Finally, the composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. 
\subsection{Explicit correlations in the non-equilibrium steady state}\label{subsection-exact}
In this section we write a formula for the stationary non-equilibrium steady state correlations between $m$- points of the chain, each of them occupied by a non-empty species, i.e. a species with label from $1$ to $N-1$. %This is motivated by the interest of understanding how actual particles of the chain are correlated among each others and also because in the steady state the holes depends on the occupation of all other species, as a consequence of the constraint $n_{N}^{x}=1-\sum_{b=1}^{N-1}n_{b}^{x}$ for arbitrary $x\in \{1,\ldots,L\}$.
 Consider $m\in \{1,\ldots,L\}$ points with coordinates given by $(x_{1},\ldots,x_{m})$, such that $x_{k}<x_{k+1}$ $\forall k\in \{1,\ldots,m-1\}$. In each of these points we assume that there is a particle of species $a_{k}\in \{1,\ldots,N-1\}$. Then, with respect to the non-equilibrium stationary distribution, the correlations among these particles are given by
\begin{equation}\label{exactCorrelations}
	\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]=\sum_{c_{1},\ldots,c_{m}=0}^{1}
	f(L,c_{1},\ldots,c_{m})\prod_{k=1}^{m}(\alpha_{a_{k}}-\beta_{a_{k}})^{c_{k}}\beta_{a_{k}}^{1-c_{k}}g_{k}(L,y_{k},c_{k},\ldots,c_{m})
\begin{equation}\label{powerCoeffSpecies}
	g_{k}(L,x_{k},c_{k},\ldots,c_{m})=\left(L+2-x_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}
\begin{equation}\label{powerCoeffNOspec}
	f(L,c_{1},\ldots,c_{m})=\frac{\Gamma(2+L-\sum_{a=1}^{m}c_{a})}{\Gamma(2+L)}
	Equation \eqref{exactCorrelations} is obtained from \eqref{ABS_1}.\\
	Deriving correlations from \eqref{ABS}
	\begin{equation}\label{exactCorrelation_1}
			\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{y_{k}}=s_{k}\}}\right]=&\frac{1}{(1+L)!}\sum_{c_{y_{1}},\ldots,c_{y_{m}}=0}^{1}\prod_{k=1}^{m}(\alpha_{s_{k}}-\beta_{s_{k}})^{1-c_{y_{k}}}\beta_{s_{k}}^{c_{y_{k}}}\left(2+L-y_{k}-\sum_{j=k}^{m}(1-c_{y_{j}})\right)\\&\prod_{x=1\,:\,x\notin \{y_{1},\ldots,y_{m}\}}^{L}\left(2+L-x-\sum_{k=1}^{m}\mathbbm{1}_{\{x<y_{k}\}}(1-c_{y_{k}})\right)
	In \eqref{exactCorrelations} we have the $c_{y_{k}}$ at the exponent but we remove the product over $x$.
	In \eqref{exactCorrelation_1} we have to keep the product over all $x$ and the "correlated coordinate" but we do not have the $c_{y}$ at the exponent. Which is better?
\paragraph{Absorption probabilities}
Knowing the non-equilibrium steady state correlations, we can read-off the absorption probabilities defined in \eqref{absProbabilitiesIntegrable}. They are given by
\begin{equation}\label{abs-probabilities}
	\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi})=\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(L,c_{1},\ldots,c_{m})\prod_{j=1}^{m}(-1)^{c_{j}-t_{j}}g_{j}(L,x_{j},c_{j},\ldots,c_{m})
where $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$. \\ \\
\textbf{Proof of \eqref{abs-probabilities}}: 
From \eqref{exactCorrelations} we have 
From first to second line we have used the binomial theorem and from third to fourth line we have changed the summation order. Therefore, comparing the above formula and \eqref{absProbabilitiesIntegrable}, we can read off the absorption probabilities given in \eqref{abs-probabilities}. 

>>>>

\usepackage{tikz}
\usetikzlibrary{decorations.pathreplacing}
% MACROS
% % 
% \newcommand{\cristian}[1]{{\bf (* {\color{ForestGreen}V:{} \small #1}*)}}
% \newcommand{\jorge}[1]{{\bf (* {\color{BlueViolet}G:{} \small #1}*)}}
% \newcommand{\rouven}[1]{{ (* {\color{red}{} \small #1}*)}}
\newtheorem{definition}{Definition}
\newtheorem{theorem}{Theorem}
\newtheorem{proposition}{Proposition}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{remark}{Remark}
\frenchspacing

@@

\newcommand{\KL}{\mathcal{K}}
\newcommand{\KR}{\hat{\mathcal{K}}}
\newcommand{\oa}{\mathbf{a}}
\newcommand{\dd}{\mathcal{D}_{u(x)}}
\newcommand{\oad}{\mathbf{a}^{+}}

@@

	H=H_{\text{left}}+H_{\text{bulk}}+H_{\text{right}}

@@

	H_{\text{bulk}}=\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}

@@

	H_{\text{right}}=\begin{pmatrix}

@@

%We define also 
%\begin{equation}\label{lambdaConditions}
%	(\alpha_{a}-\beta_{a})=\alpha_{a}-\beta_{a}\quad\text{with}\quad \sum_{a=1}^{N}(\alpha_{a}-\beta_{a})=0\,.

@@

the extra-site $u(x)$ connected to $x\in V$. We state the following duality result.
\begin{theorem}[Absorbing duality]
	The multi-species stirring process $(\bm{n}(t))_{t\geq 0}$ defined on the state space $\Omega$ with generator $\mathcal{L}$ defined in \eqref{Generator} is dual to a process $(\bm{\xi}(t))_{t\geq 0}$ defined on the enlarged state space $\widetilde{\Omega}$ with generator
	 \begin{equation}\label{DualGenerator}
		\widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
	\widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
The duality function is given by 
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
where we have introduced the reservoir densities
	\rho_{a}^{x}=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
\end{theorem}
We define the duality function of the open multi-species stirring process as
\begin{equation}\label{dualityElements}
	D(\bm{n},\bm{\xi})=\prod_{x\in V}\left(\frac{(\nu -\sum_{a=1}^{N-1}\xi_{a}^{x})!}{\nu!}\prod_{a=1}^{N-1}\frac{n_{a}^{x}!}{(n_{a}^{x}-\xi_{a}^{x})!}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\,\right)
where we recall the notation (see \eqref{rhox}) for the \textit{density} of the species $a\in \{1,\ldots,N-1\}$ imposed by the reservoir at site $x\in V$:
\rho_{a}^{x}=\frac{\alpha_{a}^{x}}{|\alpha^{x}|}\dt
The dual process is defined by its Markov generator, which reads
 \begin{equation}\label{DualGenerator}
    \widetilde{\mathcal{L}}=\sum_{(x,y)\in \mathcal{E}}\omega_{x,y}\mathcal{L}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{\mathcal{L}}_{x}
$\mathcal{L}_{x,y}$ is defined in \eqref{edgeGenerator} and, for any function $f:\widetilde{\Omega}\to \mathbb{R}$ 
\begin{equation}\label{siteDualGenerator}
    \widetilde{\mathcal{L}}_{x}f(\bm{\xi})=|\alpha^{x}|\sum_{a=1}^{N-1}\xi_{a}^{x}\left(f(\bm{\xi}-\bm{\delta}_{a}^{x}+\bm{\delta}_{N}^{x}+\bm{\delta}_{a}^{u(x)})-f(\bm{\xi})\right)\dt
\newline
The dual dynamics is described as follows. On one hand, the edge part $\mathcal{L}_{x,y}$ of the dual Markov generator gives rise to  the multi-species stirring dynamics on the graph. On the other hand, the site
part $\widetilde{\mathcal{L}}_{x}$ of the dual generator replaces a particle of any type $a\in\{1,\ldots,N-1\}$ at site $x$ with a particle of type $N$ and creates a particle of the same type $a$ at the extra-site $u(x)$. This last transition is performed with rate $|\alpha^{x}|\xi_{a}^{x}$. This means that eventually the dual process voids the graph, putting all the dual particles of species $\{1,\ldots,N-1\}$ in the extra-sites. In other words the extra-sites play the role of absorbing boundaries. 
\begin{remark} in the reversible situation, i.e. when $\forall x\in V$ we have $\rho_{a}^{x}=\rho_{a}$, the expectation  of the duality function  $D(\bm{n},\bm{\xi})$ with   $\bm{n}$ distributed as  $\mu_{rev} = \bigotimes_{x\in V}\text{Multinomial}\left(\twoj, \rho_{1},\ldots,\rho_{N}\right)$ is
\mathbb{E}_{\mu^{rev}}\left[D(\bm{n},\bm{\xi})\right]=\prod_{a=1}^{N-1}\left(\rho_{a}\right)^{\sum_{x\in V}\xi_{a}^{x}+\sum_{x\in V}\xi_{a}^{u(x)}}\qquad \forall \bm{n}\in \Omega,\quad\forall \bm{\xi}\in \widetilde{\Omega}\dt
\subsection{Proof of duality}
To prove duality between $(\bm{n}(t))_{t\geq 0}$ and $(\bm{\xi}(t))_{t\geq 0}$ we  show that \eqref{DualityRelation} is fulfilled.  To show this, we will use the Hamiltonians and their Lie algebraic description. Indeed, in this formalism the proof 
reduces to finding symmetries of the generator (for the bulk duality) and group like transformations (for the ``boundary duality'').
To a configuration $\bm{\xi}$ of the configuration space  \eqref{dualStateSpace} of a dual process we associate the vector
    |\bm{\xi}\rangle=\bigotimes_{x\in V}\left(|\xi_{1}^{x},\ldots,\xi_{N}^{x}\rangle\otimes |\xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}\rangle\right)\dt
The Hamiltonian of the dual process reads
\begin{equation}\label{DualHamiltonian}
    \widetilde{H}=\sum_{x,y\in \mathcal{E}}\omega_{x,y}\mathcal{H}_{x,y}+\sum_{x\in V}\Gamma_{x}\widetilde{H}_{x}
where $\mathcal{H}_{x,y}$ is the one defined in \eqref{edgeHamiltonian}, while 
\begin{equation}\label{siteDualHamiltonian}
    \widetilde{H}_{x}=|\alpha^{x}|\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{u(x)}\,E_{Na}^{x}-E_{aa}^{x}\right)\dt
Here we introduced the pair of bosonic operators $\mathbf{a},\,\mathbf{a}^{+}$ satisfying $[\mathbf{a},\mathbf{a}^{+}]=1$ and acting as
	\mathbf{a}^{+}|q\rangle=|q+1\rangle\qquad \mathbf{a}|q\rangle=q|q-1\rangle
	\langle q|\mathbf{a}^{+}=\langle q-1|\qquad \langle q|\mathbf{a}=(q+1) \langle q+1|
on a generic vector $\langle q|$ with $q\in \mathbb{N}_{0}$, so that in \eqref{siteDualHamiltonian} 
$(\mathbf{a}^{+})_{a}^{u(x)}$ denotes a copy of $\mathbf{a}^{+}$ acting on the extra-site $u(x)$ and on the species $a\in\{1,\ldots,N-1\}$. \\

@@

%\langle \xi_{1}^{u(x)},\ldots,0,\ldots,\xi_{N}^{u(x)}|(A^{+})_{a}^{u(x)}=0

@@

\left\{d_{x}\otimes \sum_{\xi_{1}^{u(x)},\ldots,\xi_{N-1}^{u(x)}=0}^{\infty}\left(\prod_{a=1}^{N-1}\left(\rho_{a}^{x}\right)^{\xi_{a}^{u(x)}}\right)\langle \xi_{1}^{u(x)},\ldots,\xi_{N}^{u(x)}|\right\}\,|\alpha^{x}|\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{u(x)}E_{Na}^{x}-E_{aa}^{x}\right)   

@@

\subsection{Duality for the integrable process on a line segment}\label{integrableChain-duality}
Here we consider the chain defined in Subsection \ref{subsection-description-process-LINE}. We recall that the geometry is a one dimensional chain with sites $\{1,\ldots,L\}$ where reservoirs are attached to $1$ and $L$. Now we assume the hard-core exclusion rule, i.e. $\nu=1$. The dynamics is described by an Hamiltonian that is written in terms of the first fundamental representation of $gl(N)$. Duality result stated in Subsection \ref{statementDualitySubsection} can be adapted. The extra-sites are denoted by $0$ and $L+1$, and they are connected with sites $1$ and $L$ of the chain respectively. The matrix $R_{x}$ defined in \eqref{Revmatrix} 
reduces to the identity, therefore the relation between the generators of the algebra and its transposed reduces to
\begin{equation}\label{transpostionPropertyFund}
	e_{ab}^T= e_{ba}\dt
\end{equation}
Therefore, duality matrix reads
\begin{equation}
	D=\mathcal{D}_{0}\otimes\prod_{x=1}^{L}\exp{\left(\sum_{a=1}^{N-1}e_{aN}^{x}\right)}\otimes \mathcal{D}_{L+1}
\end{equation}
where $\mathcal{D}_{0}$ and $\mathcal{D}_{L+1}$ are given in \eqref{dualityMatrix2}. 
The elements of this duality matrix are given by
\begin{equation}\label{ElementsDualityMatrixChain_1}
	D(\bm{\tau},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{\delta_{\tau_{x},a}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
\end{equation}
%\begin{equation}\label{ElementsDualityMatrixChain}
%	D(\bm{n},\bm{\xi})=\prod_{a=1}^{N-1}\alpha_{a}^{\xi_{a}^{0}}\left(\prod_{x=1}^{L}\prod_{a=1}^{N-1}\mathbbm{1}_{\{n_{a}^{x}\geq \xi_{a}^{x}\}}\right)\prod_{a=1}^{N-1}\beta_{a}^{\xi_{a}^{L+1}}\dt
%\end{equation}
The dual Hamiltonian is
\begin{equation}
	\widetilde{H}=\widetilde{H}_{\text{left}}+\sum_{x=1}^{L-1}\mathcal{H}_{x,x+1}+\widetilde{H}_{\text{right}}
\end{equation}
where $\mathcal{H}$ is defined in \eqref{H-corsivo} and 
\begin{equation}\label{boundary-H-dual-chain}
	\widetilde{H}_{\text{left}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{0}\,e_{Na}^{1}-e_{aa}^{1}\right)\qquad 	\widetilde{H}_{\text{right}}=\sum_{a=1}^{N-1}\left((\mathbf{a}^{+})_{a}^{L+1}\,e_{Na}^{L}-e_{aa}^{L}\right)\dt
\end{equation}
In the long-time limit the dual process voids the chain, i.e. all particles of types $\{1,\ldots,N-1\}$ are eventually absorbed at the extra-sites $0$ and $L+1$ and replaced by types $N$. We also notice that, up the the bosonic creation operators $(\mathbf{a}^{+})_{a}$, the dual boundary Hamiltonians \eqref{boundary-H-dual-chain} are triangular. This simplification will be crucial in Sections \ref{sectionIntegrabiliy}.\\

@@

	\left[X_{a},X_{b}\right]=(\alpha_{a}-\beta_{a})X_{b}-\lambda_{b}X_{a}\qquad\forall a,b=1,\ldots,N

@@

	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=(\alpha_{a}-\beta_{a})\langle V^{*}|\qquad\forall a=1,\ldots,N\co

@@

	\left(\beta_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)|V\rangle=-(\alpha_{a}-\beta_{a})|V\rangle\qquad\forall a=1,\ldots,N\dt

@@

		\sum_{a=1}^{N}	\langle V^{*}|\left(\alpha_{a}(X_{1}+\ldots+X_{N})-X_{a}\right)=0,%\sum_{a=1}^{N}(\alpha_{a}-\beta_{a})\langle V^{*}|=0
	and using  \eqref{ratesConditions}.

@@

 The first transformation $S_{1}$ is closely related to the duality matrix and we will see that $H^{'}$ is $\widetilde{H}^{T}$, up to the extra-site term described by the bosonic creation operator in \eqref{siteDualHamiltonian}. Therefore, $H^{'}$ is not stochastic. The second transformation $S_{2}$ diagonalizes the right boundary. Finally, the composition  $S_{2} S_{1}$ allows to simplify the commutators of the MPA algebra. 
\subsection{Explicit correlations in the non-equilibrium steady state}\label{subsection-exact}
In this section we write a formula for the stationary non-equilibrium steady state correlations between $m$- points of the chain, each of them occupied by a non-empty species, i.e. a species with label from $1$ to $N-1$.
\begin{theorem}\label{thm-correlations}[Explicit non-equilibrium steady state correlations]
Consider $m\in \{1,\ldots,L\}$ points with coordinates given by $(x_{1},\ldots,x_{m})$, such that $x_{k}<x_{k+1}$ $\forall k\in \{1,\ldots,m-1\}$. We assume that in each of these points there is a particle of species $a_{k}\in \{1,\ldots,N-1\}$. Then the $m$-point correlations with respect to the non-equilibrium steady state measure $\mu^{NESS}$ are given by 
\begin{equation}\label{exactCorrelations}
	\begin{split}
		&\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]\\=&\sum_{t_{1},\ldots,t_{m}=0}^{1}\left(\prod_{k=1}^{m}\alpha_{a_{k}}^{t_{k}}\beta_{a_{k}}^{1-t_{k}}\right)\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(L,c_{1},\ldots,c_{m})\left(\prod_{k=1}^{m}(-1)^{c_{k}-t_{k}}g_{k}(L,x_{k},c_{k},\ldots,c_{m})\right)
		\end{split}
\end{equation}
where
\begin{equation}\label{powerCoeffSpecies}
	g_{k}(L,x_{k},c_{k},\ldots,c_{m})=\left(L+2-x_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}
\end{equation}
and 
\begin{equation}\label{powerCoeffNOspec}
	f(L,c_{1},\ldots,c_{m})=\frac{\Gamma(2+L-\sum_{a=1}^{m}c_{a})}{\Gamma(2+L)}
\end{equation}
Moreover, the absorption probabilities defined in \eqref{absProbabilitiesIntegrable} are
\begin{equation}\label{abs-probabilities}
	\mathcal{P}(t_{1},\ldots,t_{m},\bm{\xi})=\sum_{c_{1}=t_{1}}^{1}\ldots\sum_{c_{m}=t_{m}}^{1}f(L,c_{1},\ldots,c_{m})\prod_{j=1}^{m}(-1)^{c_{j}-t_{j}}g_{j}(L,x_{j},c_{j},\ldots,c_{m})
\end{equation}
where $\bm{\xi}=\sum_{k=1}^{m}\delta_{a_{k}}^{x_{k}}$.
\end{theorem}
\begin{comment}
 Consider . . Then, with respect to the non-equilibrium stationary distribution, the correlations among these particles are given by
\begin{equation}\label{exactCorrelations}
	\mathbb{E}_{\mu^{\text{NESS}}}\left[\prod_{k=1}^{m}\mathbbm{1}_{\{\tau_{x_{k}}=a_{k}\}}\right]=\sum_{c_{1},\ldots,c_{m}=0}^{1}
	f(L,c_{1},\ldots,c_{m})\prod_{k=1}^{m}(\alpha_{a_{k}}-\beta_{a_{k}})^{c_{k}}\beta_{a_{k}}^{1-c_{k}}g_{k}(L,y_{k},c_{k},\ldots,c_{m})
\end{equation}
\end{comment}
\begin{comment}
where 
\begin{equation}\label{powerCoeffSpecies}
	g_{k}(L,x_{k},c_{k},\ldots,c_{m})=\left(L+2-x_{k}-\sum_{j=k}^{m}c_{j}\right)^{c_{k}}
\end{equation}
and 
\begin{equation}\label{powerCoeffNOspec}
	f(L,c_{1},\ldots,c_{m})=\frac{\Gamma(2+L-\sum_{a=1}^{m}c_{a})}{\Gamma(2+L)}
\end{equation}
\end{comment}
